---
title: "Exploratory description of the Trajetorias dataset"
author: "Juan Salamanca"
date: "2024-19-24"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 10)
```


```{r}
# install.packages("collapsibleTree")
library(collapsibleTree)
library(ggplot2)
#install.packages("geojsonsf")
library(geojsonsf)
library(sf)
library(dplyr)

```


Reference: Rorato, A.C., Dal’Asta, A.P., Lana, R.M. et al. Trajetorias: a dataset of environmental, epidemiological, and economic indicators for the Brazilian Amazon. Sci Data 10, 65 (2023). https://doi.org/10.1038/s41597-023-01962-1. 

Data source: https://zenodo.org/records/7098053. 

---


### Structure of the dataset

The Trajetorias dataset contains data of each municipality of Brazil's Legal Amazon in 2006 and 2017. It contains a rich set of indicators that allow analyzing the spatial and temporal relationship between economic trajectories, availability of natural resources and disease transmission [^1]. 

* The dataset includes **environmental indicators**, calculated from satellite images, that characterize biodiversity loss, land use and cover patterns, transport network density and climate anomalies. 

* There are **epidemiological indicators** of leishmaniasis, malaria, Chagas disease, and dengue. These infectious diseases are indicative of social and environmental vulnerability, which include poverty, lack of basic sanitation, and reduced availability of drinking water. 

* The **socioeconomic indicators**, are rural and urban populations.

* The **economic indicators**, characterize the multidimensional poverty index of rural and urban populations.

  

This is the overall structure of the dataset as described in the file *TRAJETORIAS_DATASET_Data_Source.csv*

```{r, main_indicators}
Main_indicators <- read.csv("../Trajetorias_dataset/TRAJETORIAS_DATASET_Data_Source.csv", sep=";")

collapsibleTree(
  Main_indicators,
  hierarchy = c("Dimension", "Subdimension", "Indicators"),
  width = 800,
  zoomable = FALSE,
  collapsed = FALSE
)
```

---

## Environmental dimension

The environmental dimension is divided four subdimensions: Habitat loss, Land use and cover, Climate anomalies, and Transport network density. Each subdimension has several indicators. file:TRAJETORIAS_DATASET_Environmental_dimension_indicators.csv

```{r, environment_data}
Environmental_Meta <- read.csv("../Trajetorias_dataset/TRAJETORIAS_DATASET_Environmental_dimension_indicators_METADATA.csv", sep=";")

Environmental_Meta <- na.omit(Environmental_Meta)

collapsibleTree(
  Environmental_Meta,
  hierarchy = c("Subdimension", "Indicator", "ID"),
  width = 800,
  zoomable = FALSE,
  collapsed = FALSE
)

Environmental <- read.csv("~/Documents/GitHub/Trajetorias_Brazil/Trajetorias_dataset/TRAJETORIAS_DATASET_Environmental_dimension_indicators.csv")


# Brazil map
AmazonBrazilGeo <- geojson_sf("~/Documents/GitHub/Trajetorias_Brazil/Brazilian_Legal_Amazon/Brazil_Amazon.geojson")

# Convert the GECODIGO column to integer
AmazonBrazilGeo$GEOCODIGO <- as.integer(AmazonBrazilGeo$GEOCODIGO)

# Join the environmental indicators with the geospatial data
AmazonBrazil_Environment <- left_join(AmazonBrazilGeo,Environmental,by=c( 'GEOCODIGO' ='geocode'))
```

```{r, colors}
palette <- list("Habitat loss" = list("base" = list("fill"= "seashell", "outline" = "seashell3"),"indicators"= list(
                  "Deforestation" =	"blueviolet",
                  "Forest degradation"=	"indianred4",
                  "Fires"=	"darkorange2",
                  "Mining" =	"midnightblue",
                  "Vegetation fragmentation"= "forestgreen")),
                "Land Use and Land Cover" = list("base" = list("fill"= "lavender", "outline" = "lavenderblush3"),"indicators"= list(
                  "Remnant forest" = "khaki4", 
                  "Secondary vegetation" = "darkolivegreen", 
                  "Pasture" = "darkgoldenrod4", 
                  "Crop" = "chartreuse4", 
                  "Urban area" = "deeppink4")),
                "Transportation networks" = list("base" = list("fill"= "#FFEFFF", "outline" = "gray55"),"indicators"= list(
                  "Roads network" = "darkorchid4",
                  "Waterways network" = "hotpink4",
                  "Ports" = "cadetblue")),
                "Climate anomalies" = list("base" = list("fill"= "azure", "outline" = "azure3"),"indicators"= list(
                  "Precipitation"= "steelblue",
                  "Minimum Temperature" = "brown4")),
                "Occurrence of diseases" = list("base" = list("fill"= "lightcyan", "outline" = "lightcyan3"),"indicators"= list(
                  "Cutaneous leishmaniasis" = "grey90",
                  "Visceral Leishmaniasis" = "grey90", 
                  "Malaria" = "grey90", 
                  "Chagas" = "grey90", 
                  "Dengue" = "grey90")),
                "Population" = list("base" = list("fill"= "#e3fff6", "outline" = "#9bebd2"),"indicators"= list(
                  "Rural population" = "grey90", 
                  "Urban population" = "grey90")),
                "Poverty indices" = list("base" = list("fill"= "#e6f0ff", "outline" = "#c3d0f7"),"indicators"= list(
                  "Deprivations indicators" = "grey90"))
                )
              
```

<div style="page-break-after: always"></div>

#### 1. Subdimension Habitat Loss
##### Indicator: Deforestation

Data source: PRODES (Brazilian Amazon Deforestation Monitoring Project from INPE). Available in http://terrabrasilis.dpi.inpe.br/downloads/

```{r, habitatLoss_Maps_deorg}

# Drawing the maps
AmazonBrazilMap <- ggplot(AmazonBrazil_Environment %>% filter(!is.na(period)), aes(fill = deorg)) 
# add the geometry
AmazonBrazilMap <- AmazonBrazilMap + geom_sf(color = palette$`Habitat loss`$base$outline, size = 0.1)
# set color scale
AmazonBrazilMap <- AmazonBrazilMap + scale_fill_gradient(
  low = palette$`Habitat loss`$base$fill, 
  high = palette$`Habitat loss`$indicators$Deforestation, 
  limits = c(0, 1))
# add title and subtitle
AmazonBrazilMap <- AmazonBrazilMap + labs(title = "Deforestation between 2000-2006 and 2007-2017", 
                                          subtitle = "Dimension: Environmental, Subdimension: Habitat loss, ID= deorg", 
                                          caption = "The total area of deforestation in the municipality between 2000 to 2006 and 2010 to 2016 is divided by its original* forest area", 
                                          fill = "Deforestation. km2/km2")
# facets
AmazonBrazilMap <- AmazonBrazilMap + facet_wrap(~period, drop = TRUE) 
# style
AmazonBrazilMap <- AmazonBrazilMap+ theme(plot.caption = element_text(size = 8),
                                          axis.text.x = element_blank(), 
                                          axis.text.y = element_blank(), 
                                          axis.ticks = element_blank(), 
                                          axis.title.x = element_blank(), 
                                          axis.title.y = element_blank(),
                                          panel.background = element_rect(fill = "transparent"),
                                          legend.position = "bottom")
# plot map
AmazonBrazilMap
```

```{r, habitatLoss_Maps_defor}
# Drawing the maps
AmazonBrazilMap <- ggplot(AmazonBrazil_Environment %>% filter(!is.na(period)), aes(fill = defor)) 
# add the geometry
AmazonBrazilMap <- AmazonBrazilMap + geom_sf(color = palette$`Habitat loss`$base$outline, size = 0.1)
# set color scale
AmazonBrazilMap <- AmazonBrazilMap + scale_fill_gradient(
  low = palette$`Habitat loss`$base$fill, 
  high = palette$`Habitat loss`$indicators$Deforestation, 
  limits = c(0, 1))
# add title and subtitle
AmazonBrazilMap <- AmazonBrazilMap + labs(title = "Deforestation between 2000-2006 and 2007-2017 relative to 2000", 
                                          subtitle = "Dimension: Environmental, Subdimension: Habitat loss, ID defor",
                                          caption = "The total area of deforestation in each municipality between 2000 to 2006 and 2010 to 2016 is divided by its remaining forest area in 2000", 
                                          fill = "Deforestation relative to 2000. km2/km2")
# facets
AmazonBrazilMap <- AmazonBrazilMap + facet_wrap(~period, drop = TRUE) 
# style
AmazonBrazilMap <- AmazonBrazilMap+ theme(plot.caption = element_text(size = 8),axis.text.x = element_blank(), 
  axis.text.y = element_blank(), axis.ticks = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank(),                                           panel.background = element_rect(fill = "transparent"),legend.position = "bottom")
# plot map
AmazonBrazilMap
```

##### Indicator: Forest degradation

Data source: DEGRAD and DETER (Real-Time Deforestation Detection System) programs from INPE. Available in http://www.obt.inpe.br/OBT/assuntos/programas/amazonia/degrad/acesso-ao-dados-do-degrad http://terrabrasilis.dpi.inpe.br/downloads/

```{r, habitatLoss_Maps_dgorg}

# Drawing the maps
AmazonBrazilMap <- ggplot(AmazonBrazil_Environment %>% filter(!is.na(period)), aes(fill = dgorg)) 
# add the geometry
AmazonBrazilMap <- AmazonBrazilMap + geom_sf(color = palette$`Habitat loss`$base$outline, size = 0.1)
# set color scale
AmazonBrazilMap <- AmazonBrazilMap + scale_fill_gradient(
  low = palette$`Habitat loss`$base$fill, 
  high = palette$`Habitat loss`$indicators$`Forest degradation`, 
  limits = c(0, 1))
# add title and subtitle
AmazonBrazilMap <- AmazonBrazilMap + labs(title = "Forest degradation between 2007-2017", 
                                          subtitle = "Dimension: Environmental, Subdimension: Habitat loss, ID: dgorg",
                                          caption = "The total area of forest degradation in the municipality between 2007 to 2017 is divided by its original* forest area.", 
                                          fill = "Forest degradation. km2/km2")
# facets
#AmazonBrazilMap <- AmazonBrazilMap + facet_wrap(~period, drop = TRUE) 
# style
AmazonBrazilMap <- AmazonBrazilMap+ theme(plot.caption = element_text(size = 8),axis.text.x = element_blank(), 
  axis.text.y = element_blank(), axis.ticks = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank(),                                           panel.background = element_rect(fill = "transparent"),legend.position = "bottom")
# plot map
AmazonBrazilMap
```

```{r, habitatLoss_Maps_dgfor}

# Drawing the maps
AmazonBrazilMap <- ggplot(AmazonBrazil_Environment %>% filter(!is.na(period)), aes(fill = dgfor)) 
# add the geometry
AmazonBrazilMap <- AmazonBrazilMap + geom_sf(color = palette$`Habitat loss`$base$outline, size = 0.1)
# set color scale
AmazonBrazilMap <- AmazonBrazilMap + scale_fill_gradient(
  low = palette$`Habitat loss`$base$fill, 
  high = palette$`Habitat loss`$indicators$`Forest degradation`,
  limits = c(0, 1))
# add title and subtitle
AmazonBrazilMap <- AmazonBrazilMap + labs(title = "Forest degradation between 2007-2017 relative to 2007", 
                                          subtitle = "Dimension: Environmental, Subdimension: Habitat loss",
                                          caption = "	the total area of forest degradation between 2007 to 2017 is divided by its area with remaining forest in 2007", 
                                          fill = "Forest degradation relative to 2007. km2/km2")
# facets
#AmazonBrazilMap <- AmazonBrazilMap + facet_wrap(~period, drop = TRUE) 
# style
AmazonBrazilMap <- AmazonBrazilMap+ theme(plot.caption = element_text(size = 8),axis.text.x = element_blank(), 
  axis.text.y = element_blank(), axis.ticks = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank(),                                           panel.background = element_rect(fill = "transparent"),legend.position = "bottom")
# plot map
AmazonBrazilMap
```

##### Indicator: Fires

Data source: MODIS Burned Area Product. Available at https://modis-fire.umd.edu/ba.html

```{r, habitatLoss_Maps_fire}

# Drawing the maps
AmazonBrazilMap <- ggplot(AmazonBrazil_Environment%>% filter(!is.na(period)), aes(fill = fire)) 
# add the geometry
AmazonBrazilMap <- AmazonBrazilMap + geom_sf(color = palette$`Habitat loss`$base$outline, size = 0.1)
# set color scale
AmazonBrazilMap <- AmazonBrazilMap + scale_fill_gradient(
  low = palette$`Habitat loss`$base$fill, 
  high = palette$`Habitat loss`$indicators$Fires, 
  limits = c(0, 1))
# add title and subtitle
AmazonBrazilMap <- AmazonBrazilMap + labs(title = "Municipalities with burning events 2006 and 2017", 
                                          subtitle = "Dimension: Environmental, Subdimension: Habitat loss, ID: fire",
                                          caption = "Proportion of the municipality area that experienced at least one burning event between 2001 and 2006 and 2012 and 2017",
                                          fill = "Proportion of burning events. km2/km2")
# facets
AmazonBrazilMap <- AmazonBrazilMap + facet_wrap(~period, drop = TRUE) 
# style
AmazonBrazilMap <- AmazonBrazilMap+ theme(plot.caption = element_text(size = 8),axis.text.x = element_blank(), 
  axis.text.y = element_blank(), axis.ticks = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank(),                                           panel.background = element_rect(fill = "transparent"),legend.position = "bottom")
# plot map
AmazonBrazilMap
```

##### Indicator: Mining

Data source: MAPBIOMAS Project (Collection 6). Available at https://mapbiomas.org/colecoes-mapbiomas-1?cama_set_language=pt-BR

```{r, habitatLoss_Maps_minig}

# Drawing the maps
AmazonBrazilMap <- ggplot(AmazonBrazil_Environment%>% filter(!is.na(period)), aes(fill = asin(sqrt(mining))) )
# add the geometry
AmazonBrazilMap <- AmazonBrazilMap + geom_sf(color = palette$`Habitat loss`$base$outline, size = 0.1)
# set color scale
AmazonBrazilMap <- AmazonBrazilMap + scale_fill_gradient(
  low = palette$`Habitat loss`$base$fill, 
  high = palette$`Habitat loss`$indicators$Mining)
# add title and subtitle
AmazonBrazilMap <- AmazonBrazilMap + labs(title = "Mining 2006 and 2017", 
                                          subtitle = "Dimension: Environmental, Subdimension: Habitat loss, ID: mining",
                                          caption = "Proportion of the municipality area used for mining (industrial and artisanal) in 2006 and 2017",
                                          fill = "Mining. km2/km2. Arcsine Square Root Transformation")
# facets
AmazonBrazilMap <- AmazonBrazilMap + facet_wrap(~period, drop = TRUE) 
# style
AmazonBrazilMap <- AmazonBrazilMap+ theme(plot.caption = element_text(size = 8),axis.text.x = element_blank(), 
  axis.text.y = element_blank(), axis.ticks = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank(),                                           panel.background = element_rect(fill = "transparent"),legend.position = "bottom")
# plot map
AmazonBrazilMap
```

##### Indicator: Vegetation fragmentation

Data source: Camara et al., 2020. Available at Câmara, G et al. (2020): Land use and land cover maps for Amazon biome in Brazil for 2001-2019 derived from MODIS time series (pangaea.de)

```{r, habitatLoss_Maps_fragmentation_core}

# Drawing the maps
AmazonBrazilMap <- ggplot(AmazonBrazil_Environment%>% filter(!is.na(period)), aes(fill = core))
# add the geometry
AmazonBrazilMap <- AmazonBrazilMap + geom_sf(color = palette$`Habitat loss`$base$outline, size = 0.1)
# set color scale
AmazonBrazilMap <- AmazonBrazilMap + scale_fill_gradient(
  low = palette$`Habitat loss`$base$fill, 
  high = palette$`Habitat loss`$indicators$`Vegetation fragmentation`)
# add title and subtitle
AmazonBrazilMap <- AmazonBrazilMap + labs(title = "Natural vegetation cores in 2006 and 2017", 
                                          subtitle = "Dimension: Environmental, Subdimension: Habitat loss, ID: core",
                                          caption = "Proportion of the municipality area classified as natural vegetation cores in 2006 and 2017",
                                          fill = "Natural vegetation area. km2/km2")
# facets
AmazonBrazilMap <- AmazonBrazilMap + facet_wrap(~period, drop = TRUE) 
# style
AmazonBrazilMap <- AmazonBrazilMap+ theme(plot.caption = element_text(size = 8),axis.text.x = element_blank(), 
  axis.text.y = element_blank(), axis.ticks = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank(),                                           panel.background = element_rect(fill = "transparent"),legend.position = "bottom")
# plot map
AmazonBrazilMap
```

Data source: Camara et al., 2020. Available at Câmara, G et al. (2020): Land use and land cover maps for Amazon biome in Brazil for 2001-2019 derived from MODIS time series (pangaea.de)

```{r, habitatLoss_Maps_fragmentation_edge}

# Drawing the maps
AmazonBrazilMap <- ggplot(AmazonBrazil_Environment%>% filter(!is.na(period)), aes(fill = edge))
# add the geometry
AmazonBrazilMap <- AmazonBrazilMap + geom_sf(color = palette$`Habitat loss`$base$outline, size = 0.1)
# set color scale
AmazonBrazilMap <- AmazonBrazilMap + scale_fill_gradient(
  low = palette$`Habitat loss`$base$fill, 
  high = palette$`Habitat loss`$indicators$`Vegetation fragmentation`)
# add title and subtitle
AmazonBrazilMap <- AmazonBrazilMap + labs(title = "Perimeter of natural vegetation 2006 and 2017", 
                                          subtitle = "Dimension: Environmental, Subdimension: Habitat loss, ID: edge",
                                          caption = "Ratio between the total edge length (perimeter) of the natural vegetation patches within the municipality and the square rooted total area of these patches, in 2006 and 2017",
                                          fill = "Perimeter of natural vegetation area. m/sqrt(m2)")
# facets
AmazonBrazilMap <- AmazonBrazilMap + facet_wrap(~period, drop = TRUE) 
# style
AmazonBrazilMap <- AmazonBrazilMap+ theme(plot.caption = element_text(size = 8),axis.text.x = element_blank(), 
  axis.text.y = element_blank(), axis.ticks = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank(),                                           panel.background = element_rect(fill = "transparent"),legend.position = "bottom")
# plot map
AmazonBrazilMap
```

---



#### 2. Subdimension Land Use and Land Recovery
##### Indicator: Remnant forest

Data source: PRODES (Brazilian Amazon Deforestation Monitoring Project from INPE). Available in http://terrabrasilis.dpi.inpe.br/downloads/

```{r, landUse_Maps_refor}

# Drawing the maps
AmazonBrazilMap <- ggplot(AmazonBrazil_Environment %>% filter(!is.na(period)), aes(fill = refor)) 
# add the geometry
AmazonBrazilMap <- AmazonBrazilMap + geom_sf(color = palette$`Land Use and Land Cover`$base$outline, size = 0.1)
# set color scale
AmazonBrazilMap <- AmazonBrazilMap + scale_fill_gradient(
  low = palette$`Land Use and Land Cover`$base$fill, 
  high = palette$`Land Use and Land Cover`$indicators$`Remnant forest`, 
  limits = c(0, 1))
# add title and subtitle
AmazonBrazilMap <- AmazonBrazilMap + labs(title = "Remnant forest in 2007 and 2017", 
                                          subtitle = "Dimension: Environmental, Subdimension: Land Use and Land Cover, ID= refor", 
                                          caption = "The remnant forest area in 2007 and 2017 were divided by the original* forest area in the municipality", 
                                          fill = "Remnant forest. km2/km2")
# facets
AmazonBrazilMap <- AmazonBrazilMap + facet_wrap(~period, drop = TRUE) 
# style
AmazonBrazilMap <- AmazonBrazilMap+ theme(plot.caption = element_text(size = 8),axis.text.x = element_blank(), 
  axis.text.y = element_blank(), axis.ticks = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank(),                                           panel.background = element_rect(fill = "transparent"),legend.position = "bottom")
# plot map
AmazonBrazilMap
```

##### Indicator: Secondary vegetation

Data source: Camara et al., 2020. Available at Câmara, G et al. (2020): Land use and land cover maps for Amazon biome in Brazil for 2001-2019 derived from MODIS time series (pangaea.de)


```{r, landUse_Maps_secveg}

# Drawing the maps
AmazonBrazilMap <- ggplot(AmazonBrazil_Environment %>% filter(!is.na(period)), aes(fill = secveg)) 
# add the geometry
AmazonBrazilMap <- AmazonBrazilMap + geom_sf(color = palette$`Land Use and Land Cover`$base$outline, size = 0.1)
# set color scale
AmazonBrazilMap <- AmazonBrazilMap + scale_fill_gradient(
  low = palette$`Land Use and Land Cover`$base$fill, 
  high = palette$`Land Use and Land Cover`$indicators$`Secondary vegetation`, 
  limits = c(0, 1))
# add title and subtitle
AmazonBrazilMap <- AmazonBrazilMap + labs(title = "Secondary vegetation in 2007 and 2017", 
                                          subtitle = "Dimension: Environmental, Subdimension: Land Use and Land Cover, ID= secveg", 
                                          caption = "Proportion of the municipality area covered with secondary vegetation in 2006 and 2017
", 
                                          fill = "Secondary vegetation. km2/km2")
# facets
AmazonBrazilMap <- AmazonBrazilMap + facet_wrap(~period, drop = TRUE) 
# style
AmazonBrazilMap <- AmazonBrazilMap+ theme(plot.caption = element_text(size = 8),axis.text.x = element_blank(), 
  axis.text.y = element_blank(), axis.ticks = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank(),                                           panel.background = element_rect(fill = "transparent"),legend.position = "bottom")
# plot map
AmazonBrazilMap
```

##### Indicator: Pasture

Data source: Camara et al., 2020. Available at Câmara, G et al. (2020): Land use and land cover maps for Amazon biome in Brazil for 2001-2019 derived from MODIS time series (pangaea.de)


```{r, landUse_Maps_pasture}

# Drawing the maps
AmazonBrazilMap <- ggplot(AmazonBrazil_Environment %>% filter(!is.na(period)), aes(fill = pasture)) 
# add the geometry
AmazonBrazilMap <- AmazonBrazilMap + geom_sf(color = palette$`Land Use and Land Cover`$base$outline, size = 0.1)
# set color scale
AmazonBrazilMap <- AmazonBrazilMap + scale_fill_gradient(
  low = palette$`Land Use and Land Cover`$base$fill, 
  high = palette$`Land Use and Land Cover`$indicators$Pasture, 
  limits = c(0, 1))
# add title and subtitle
AmazonBrazilMap <- AmazonBrazilMap + labs(title = "Pasture areas in 2006 and 2017", 
                                          subtitle = "Dimension: Environmental, Subdimension: Land Use and Land Cover, ID= pasture", 
                                          caption = "Proportion of the municipality area classified as pasture in 2006 and 2017
", 
                                          fill = "Pasture area. km2/km2")
# facets
AmazonBrazilMap <- AmazonBrazilMap + facet_wrap(~period, drop = TRUE) 
# style
AmazonBrazilMap <- AmazonBrazilMap+ theme(plot.caption = element_text(size = 8),axis.text.x = element_blank(), 
  axis.text.y = element_blank(), axis.ticks = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank(),                                           panel.background = element_rect(fill = "transparent"),legend.position = "bottom")
# plot map
AmazonBrazilMap
```

##### Indicator: Crop

Data source: Camara et al., 2020. Available at Câmara, G et al. (2020): Land use and land cover maps for Amazon biome in Brazil for 2001-2019 derived from MODIS time series (pangaea.de)


```{r, landUse_Maps_crop}

# Drawing the maps
AmazonBrazilMap <- ggplot(AmazonBrazil_Environment %>% filter(!is.na(period)), aes(fill = crop)) 
# add the geometry
AmazonBrazilMap <- AmazonBrazilMap + geom_sf(color = palette$`Land Use and Land Cover`$base$outline, size = 0.1)
# set color scale
AmazonBrazilMap <- AmazonBrazilMap + scale_fill_gradient(
  low = palette$`Land Use and Land Cover`$base$fill, 
  high = palette$`Land Use and Land Cover`$indicators$Crop, 
  limits = c(0, 1))
# add title and subtitle
AmazonBrazilMap <- AmazonBrazilMap + labs(title = "Crop areas in 2006 and 2017", 
                                          subtitle = "Dimension: Environmental, Subdimension: Land Use and Land Cover, ID= pasture", 
                                          caption = "Proportion of the municipality area classified as crop in 2006 and 2017", 
                                          fill = "Crop areas. km2/km2")
# facets
AmazonBrazilMap <- AmazonBrazilMap + facet_wrap(~period, drop = TRUE) 
# style
AmazonBrazilMap <- AmazonBrazilMap+ theme(plot.caption = element_text(size = 8),axis.text.x = element_blank(), 
  axis.text.y = element_blank(), axis.ticks = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank(),                                           panel.background = element_rect(fill = "transparent"),legend.position = "bottom")
# plot map
AmazonBrazilMap
```

##### Indicator: Urban area

Data source: MAPBIOMAS Project (Collection 6)
Available at https://mapbiomas.org/colecoes-mapbiomas-1?cama_set_language=pt-BR


```{r, landUse_Maps_urban}

# Drawing the maps
AmazonBrazilMap <- ggplot(AmazonBrazil_Environment %>% filter(!is.na(period)), aes(fill = asin(sqrt(urban)))) 
# add the geometry
AmazonBrazilMap <- AmazonBrazilMap + geom_sf(color = palette$`Land Use and Land Cover`$base$outline, size = 0.1)
# set color scale
AmazonBrazilMap <- AmazonBrazilMap + scale_fill_gradient(
  low = palette$`Land Use and Land Cover`$base$fill, 
  high = palette$`Land Use and Land Cover`$indicators$`Urban area`, 
  limits = c(0, 1))
# add title and subtitle
AmazonBrazilMap <- AmazonBrazilMap + labs(title = "Urban area in 2006 and 2017", 
                                          subtitle = "Dimension: Environmental, Subdimension: Land Use and Land Cover, ID= urban", 
                                          caption = "Proportion of the municipality area classified as urban in 2006 and 2017", 
                                          fill = "Urban area. Arcsine Squared root transformation. km2/km2")
# facets
AmazonBrazilMap <- AmazonBrazilMap + facet_wrap(~period, drop = TRUE) 
# style
AmazonBrazilMap <- AmazonBrazilMap+ theme(plot.caption = element_text(size = 8),axis.text.x = element_blank(), 
  axis.text.y = element_blank(), axis.ticks = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank(),                                           panel.background = element_rect(fill = "transparent"),legend.position = "bottom")
# plot map
AmazonBrazilMap
```

---


#### 3. Transportation Networks
##### Indicator: Roads networks

Data source: LAPIG (Laboratory of Image Processing and Geoprocessing at the Federal University of Goiás) and RAISG (Amazon Network of Georeferenced Social and Environmental Information). Available at Mapas y Publicaciones – RAISG (amazoniasocioambiental.org)


```{r, transportation_Maps_road}

# Drawing the maps
AmazonBrazilMap <- ggplot(AmazonBrazil_Environment %>% filter(!is.na(period)), aes(fill = road)) 
# add the geometry
AmazonBrazilMap <- AmazonBrazilMap + geom_sf(color = palette$`Transportation networks`$base$outline, size = 0.1)
# set color scale
AmazonBrazilMap <- AmazonBrazilMap + scale_fill_gradient(
  low = palette$`Transportation networks`$base$fill, 
  high = palette$`Transportation networks`$indicators$`Roads network`, 
  limits = c(0, 1))
# add title and subtitle
AmazonBrazilMap <- AmazonBrazilMap + labs(title = "Roads in 2007 and 2017", 
                                          subtitle = "Dimension: Environmental, Subdimension: Transportation Networks, ID= road", 
                                          caption = "The total length of roads within the municipality divided by the square root of its area, in 2010 and 2017
", 
                                          fill = "Roads. m/sqrt(m2)
")
# facets
AmazonBrazilMap <- AmazonBrazilMap + facet_wrap(~period, drop = TRUE) 
# style
AmazonBrazilMap <- AmazonBrazilMap+ theme(plot.caption = element_text(size = 8),axis.text.x = element_blank(), 
  axis.text.y = element_blank(), axis.ticks = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank(),                                           panel.background = element_rect(fill = "transparent"),legend.position = "bottom")
# plot map
AmazonBrazilMap
```

##### Indicator: Waterways network

Data source: DNIT (National Infrastructure and Transport Department) and ANTAQ (National Water Transport Agency). Available at https://metadados.snirh.gov.br/geonetwork/srv/api/records/48e26e99-db01-45dc-a270-79f27680167b

```{r, transportation_Maps_river}

# Drawing the maps
AmazonBrazilMap <- ggplot(AmazonBrazil_Environment %>% filter(!is.na(period)), aes(fill = river)) 
# add the geometry
AmazonBrazilMap <- AmazonBrazilMap + geom_sf(color = palette$`Transportation networks`$base$outline, size = 0.1)
# set color scale
AmazonBrazilMap <- AmazonBrazilMap + scale_fill_gradient(
  low = palette$`Transportation networks`$base$fill, 
  high = palette$`Transportation networks`$indicators$`Waterways network`, 
  limits = c(0, 1))
# add title and subtitle
AmazonBrazilMap <- AmazonBrazilMap + labs(title = "Waterways in 2010", 
                                          subtitle = "Dimension: Environmental, Subdimension: Transportation Networks, ID= river", 
                                          caption = "The total length of waterways (both main or inexpressive rivers) within the municipality divided by the square root of its area", 
                                          fill = "Roads. m/sqrt(m2)
")
# facets
#AmazonBrazilMap <- AmazonBrazilMap + facet_wrap(~period, drop = TRUE) 
# style
AmazonBrazilMap <- AmazonBrazilMap+ theme(plot.caption = element_text(size = 8),axis.text.x = element_blank(), 
  axis.text.y = element_blank(), axis.ticks = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank(),                                           panel.background = element_rect(fill = "transparent"),legend.position = "bottom")
# plot map
AmazonBrazilMap
```
---

##### Indicator: Ports

Data source: DNIT (National Infrastructure and Transport Department). Available at https://servicos.dnit.gov.br/vgeo/


```{r, transportation_Maps_port}

# Drawing the maps
AmazonBrazilMap <- ggplot(AmazonBrazil_Environment %>% filter(!is.na(period)), aes(fill = port)) 
# add the geometry
AmazonBrazilMap <- AmazonBrazilMap + geom_sf(color = palette$`Transportation networks`$base$outline, size = 0.1)
# set color scale
AmazonBrazilMap <- AmazonBrazilMap + scale_fill_gradient(
  low = palette$`Transportation networks`$base$fill, 
  high = palette$`Transportation networks`$indicators$Ports)
# add title and subtitle
AmazonBrazilMap <- AmazonBrazilMap + labs(title = "Number of ports in 2010", 
                                          subtitle = "Dimension: Environmental, Subdimension: Transportation Networks, ID= port", 
                                          caption = "Number of ports", 
                                          fill = "Ports
")
# facets
#AmazonBrazilMap <- AmazonBrazilMap + facet_wrap(~period, drop = TRUE) 
# style
AmazonBrazilMap <- AmazonBrazilMap+ theme(plot.caption = element_text(size = 8),axis.text.x = element_blank(), 
  axis.text.y = element_blank(), axis.ticks = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank(),                                           panel.background = element_rect(fill = "transparent"),legend.position = "bottom")
# plot map
AmazonBrazilMap
```
---



---



#### 4. Climate Anomalies
##### Indicator: Precipitation

Data source: WorldClim - Historical monthly weather data
Available at https://www.worldclim.org/data/monthlywth.html#

```{r, climate_Maps_precp}

# Drawing the maps
AmazonBrazilMap <- ggplot(AmazonBrazil_Environment %>% filter(!is.na(period)), aes(fill = precp)) 
# add the geometry
AmazonBrazilMap <- AmazonBrazilMap + geom_sf(color = palette$`Climate anomalies`$base$outline, size = 0.1)
# set color scale
AmazonBrazilMap <- AmazonBrazilMap + scale_fill_gradient(
  low = palette$`Climate anomalies`$base$fill, 
  high = palette$`Climate anomalies`$indicators$Precipitation, 
  limits = c(0, 1))
# add title and subtitle
AmazonBrazilMap <- AmazonBrazilMap + labs(title = "Mean positive precipitation anomaly area bewteen 1996-2006 and 2007-2017", 
                                          subtitle = "Dimension: Environmental, Subdimension: Climate Anomalies, ID= precp", 
                                          caption = "Mean positive precipitation anomaly area observed during the dry seasons from 1996-2006 and 2007-2017, divided by the total area", 
                                          fill = "Precipitation area. m2/m2")
# facets
AmazonBrazilMap <- AmazonBrazilMap + facet_wrap(~period, drop = TRUE) 
# style
AmazonBrazilMap <- AmazonBrazilMap+ theme(plot.caption = element_text(size = 8),axis.text.x = element_blank(), 
  axis.text.y = element_blank(), axis.ticks = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank(),                                           panel.background = element_rect(fill = "transparent"),legend.position = "bottom")
# plot map
AmazonBrazilMap
```

```{r, climate_Maps_precn}

# Drawing the maps
AmazonBrazilMap <- ggplot(AmazonBrazil_Environment %>% filter(!is.na(period)), aes(fill = precn)) 
# add the geometry
AmazonBrazilMap <- AmazonBrazilMap + geom_sf(color = palette$`Climate anomalies`$base$outline, size = 0.1)
# set color scale
AmazonBrazilMap <- AmazonBrazilMap + scale_fill_gradient(
  low = palette$`Climate anomalies`$base$fill, 
  high = palette$`Climate anomalies`$indicators$Precipitation, 
  limits = c(0, 1))
# add title and subtitle
AmazonBrazilMap <- AmazonBrazilMap + labs(title = "Mean negative precipitation anomaly area bewteen 1996-2006 and 2007-2017", 
                                          subtitle = "Dimension: Environmental, Subdimension: Climate Anomalies, ID= precn", 
                                          caption = "Mean negative precipitation anomaly area observed during the dry seasons from 1996-2006
 and 2007-2017, divided by the total area", 
                                          fill = "Precipitation area. m2/m2")
# facets
AmazonBrazilMap <- AmazonBrazilMap + facet_wrap(~period, drop = TRUE) 
# style
AmazonBrazilMap <- AmazonBrazilMap+ theme(plot.caption = element_text(size = 8),axis.text.x = element_blank(), 
  axis.text.y = element_blank(), axis.ticks = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank(),                                           panel.background = element_rect(fill = "transparent"),legend.position = "bottom")
# plot map
AmazonBrazilMap
```

##### Indicator: Temperature

Data source: WorldClim - Historical monthly weather data
Available at https://www.worldclim.org/data/monthlywth.html#

```{r, climate_Maps_tempp}

# Drawing the maps
AmazonBrazilMap <- ggplot(AmazonBrazil_Environment %>% filter(!is.na(period)), aes(fill = tempp)) 
# add the geometry
AmazonBrazilMap <- AmazonBrazilMap + geom_sf(color = palette$`Climate anomalies`$base$outline, size = 0.1)
# set color scale
AmazonBrazilMap <- AmazonBrazilMap + scale_fill_gradient(
  low = palette$`Climate anomalies`$base$fill, 
  high = palette$`Climate anomalies`$indicators$`Minimum Temperature`, 
  limits = c(0, 1))
# add title and subtitle
AmazonBrazilMap <- AmazonBrazilMap + labs(title = "Mean positive temperature anomaly area bewteen 1996-2006 and 2007-2017", 
                                          subtitle = "Dimension: Environmental, Subdimension: Climate Anomalies, ID= tempp", 
                                          caption = "Mean positive temperature anomaly area observed during the cold seasons from 1996-2006 and 2007-2017, divided by the total area", 
                                          fill = "Temperature area. m2/m2")
# facets
AmazonBrazilMap <- AmazonBrazilMap + facet_wrap(~period, drop = TRUE) 
# style
AmazonBrazilMap <- AmazonBrazilMap+ theme(plot.caption = element_text(size = 8),axis.text.x = element_blank(), 
  axis.text.y = element_blank(), axis.ticks = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank(),                                           panel.background = element_rect(fill = "transparent"),legend.position = "bottom")
# plot map
AmazonBrazilMap
```




### Notes
[^1]:(Rewrote from original published at https://github.com/Trajetorias-Sinbiose/Trajetorias_dataset.)